Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic

### 2020-06-10 write-up/covid-writeup-science.tex
Create covid-writeup-science.tex

\title{The Hypothesis of Testing: Paradoxes arising out of reported coronavirus case-counts}

@@

\begin{sciabstract}
  Many statisticians, epidemiologists, economists and data scientists have registered serious reservations regarding the reported coronavirus case-counts. Limited testing capacity across the country has been  identified as a key driver of suppressed coronavirus case-counts.  The calls to increase testing capacity are well-justified as they become a more frequent point of discussion in the public sphere.  While expanded testing is a laudable goal, selection bias will impact estimates of disease prevalence and the effective reproduction number until the entire population is sampled.  Moreover, tests are imperfect as false positive/negative rates interact in complex ways with selection bias.  In this paper, we attempt to clarify this interaction.  Through simple calculations, we demonstrate pitfalls and paradoxes that can arise when considering case-count data in the presence of selection bias and measurement error.
\end{sciabstract}

>>>>

\title{Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic}

% \title{The Hypothesis of Testing: Paradoxes arising out of reported coronavirus case-counts}

@@

\begin{sciabstract}
  Coronavirus case-counts have influenced governments on their policies and decision making, while researchers have used them to drive their forecast models and predictions. Throughout this, many have also cited limited testing as the key driver of weak information on the COVID-19 pandemic. The calls to increase testing capacity are well-justified as they become a more frequent point of discussion in the public sphere. While expanded testing is a laudable goal, selection bias will impact estimates of disease prevalence and the effective reproduction number until the entire population is sampled. Moreover, tests are imperfect as false positive/negative rates interact in complex ways with selection bias. Indeed, measurement error and selection bias are the two greatest problems limiting our understanding of the current COVID-19 pandemic; neither can be fully addressed by increased testing capacity. In this paper, we attempt to clarify this point. Through simple calculations, we demonstrate pitfalls and paradoxes that can arise when considering case-count data in the presence of selection bias and measurement error.
\end{sciabstract}
